---
title: "ProductionAnalysis xlxSheeting"
author: "Brandon Mayer"
date: "9/13/2021"
output: html_document
---
```{r setup, include=FALSE,echo=FALSE,}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Walkthrough to organize .xlsx file into CSV files for analysis in "Production_Processing.R" file.

Paste the location of the Raw or remove "readline()" and type in file location. 

```{r}
# absol_path <- "C:/Users/brand/Dropbox/Project_Data/SRER/Production/SRERexl2021"
# absol_path <- ""
print("Paste the location of the .xlsx files");absol_path <- readline();print("Type the year collected");year <-readline()

# create a location in your working directory to store sheeted data
dir.create(paste0(getwd(),"/Data/"));dir.create(paste0(getwd(),"/Data/Production"))
finalfolder <-paste0(paste0(absol_path,"/",year,"_sheeted"));dir.create(finalfolder)
# folder_files <- list.files("C:/Users/brand/Dropbox/Project_Data/SRER/Production/SRERexl2020",recursive = T, include.dirs = F, pattern = ".xlsx")
(folder_files <- list.files(absol_path,recursive = F, include.dirs = F, pattern = ".xlsx"))

```

Splits xlsx into CSVs for analysis

```{r}
for (i in folder_files) {
  path_to_xlsx <- paste(absol_path,i,sep="/")
  #path_to_xlsx <- "C:/Users/brand/Dropbox/Project_Data/SRER/Production/SRERexl2021/20211001_UAE_T1.xlsx"
  metadata <- readxl::read_excel(path_to_xlsx, sheet = 1)
  Pasture <- gsub(" ","",gsub("Site:","", sub('-.*', '\\1',sub('.*>', '\\1', metadata[1,1])), fixed = TRUE))
  #Transect <- tail(unlist(stringr::str_split(metadata[1,1]," ")), n=1) 
  Transect <- stringr::str_split(gsub(".xlsx","",i), "_")[[1]][2] #use if sites are separated
  date <- gsub("-","",format(strptime(gsub(" ","",as.character(sub('.*:', '\\1', metadata[2,1]))), format="%m/%d/%Y"), "%Y-%m-%d"))
  base <- paste0(Pasture,"_",Transect)
  sheet_names <- readxl::excel_sheets(path_to_xlsx)
  # sheet_names <- readxl::excel_sheets(path_to_xlsx)[2:4]
  # base <- gsub("\\..*","",basename(path_to_xlsx))
  for (j in sheet_names) {
    # j <- sheet_names[2]
    data <- readxl::read_excel(path_to_xlsx, sheet=j)
    if (j == "Frequency (by quadrat)") {
      name <- "Freq"
    }else if (j == "Comparative Yield") {
      name <- "CY"
    }else if (j == "Point Ground Cover") {
      name <- "PGC"
    }else{
      name <- "DWR"
    }
    write.csv(data, file=paste0(finalfolder,"/",date,"_",base,"_",name,".csv"))
  }
}
list.files(finalfolder)
```

------------------------------------------------------------------------------STOP------------------------------------------------------------------------

If sites are split using multiple tablets and not merged prior to analysis use the following scrip to rbind. Ignore otherwise. 
  -   Identifies which files are from the same site. 
  -   Rbinds and rewrites files to new folder. 
```{r}
SheetedFiles<- list.files(finalfolder,full.names = T)
sites <-unique(gsub(".1\\b","", unique(unlist(lapply(stringr::str_split(gsub(".csv","",list.files(finalfolder)), "_"), FUN = `[`, 3)))))
Methods<- unique(unlist(lapply(stringr::str_split(gsub(".csv","",list.files(finalfolder)), "_"), FUN = `[`, 4)))
for (s in sites) {
  FocusedFiles <- SheetedFiles[stringr::str_detect(SheetedFiles, s)]
  for (m in Methods) {
      SiteMethodFiles <- FocusedFiles[stringr::str_detect(FocusedFiles, m)]
      date<- unique(unlist(lapply(stringr::str_split(gsub(".csv","",basename(SiteMethodFiles)), "_"), FUN = `[`, 1)))
      BoundSite <-data.table::rbindlist(lapply(SiteMethodFiles, read.csv)) %>% select(-X)
      BoundSite[order(BoundSite$Transect,BoundSite$Sample)]
      write.csv(BoundSite,file=paste0(finalfolder,"/",date,"_",s,"_",m,".csv"))
  }
}
```

