---
title: "FileClean_Separeate"
author: "Brandon Mayer"
date: "12/28/2020"
output: html_document
---
This file includes an extra step in setup for the utilization analysis process. 
This will include errors such as correcting: 
- metric systems
- wrongly entered plant names

Corrected folders will saved as csv's for final analysis 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(stringr)
```

# Upload file into R 
```{r, include=FALSE}
# Navigate to Utilization year through the "go to director" option in the files tab. 
# Paste in consule once prompted by "readline.
print("Paste below V")
path<- gsub("\\\\", "/", readline())
```
Choose a pasture folder that's been completed. This folder, and all the contents 
Will be broken down into csv's of each transect and event and stored in a processed filer 

```{r }
Completed_pasture_folder <- "12C"
```

Breaks up te folder by sheet and stores them in the cleaned_data folder 
```{r pressure, echo=FALSE}

FolderPath<-paste0(path,"/",Completed_pasture_folder) 
(file.list <- list.files(path = FolderPath, pattern='*.xlsx',full.names = T))
finalfolder<-gsub("data","Cleaned_Data",FolderPath)
dir.create(finalfolder)

for (i in file.list) {
  path_to_xlsx <-i
  # path_to_xlsx <- "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/data/2021/12C/12C1.xlsx"
  metadata <- readxl::read_excel(path_to_xlsx, sheet = 1)
  Pasture <- gsub(" ", "", sub('-.*', '\\1',sub('.*>', '\\1', metadata[1,1])), fixed = TRUE)
  # When processing 2S check into adding date to each file name
  #                                                                                               Look---V
  # date <- format(strptime(as.character(gsub(" ","",sub('.*:', '\\1', metadata[2,2]))), "%m/%d/%Y"), "%Y%m%d")
  
  Transect <- gsub(" ", "", gsub(Pasture,"",str_replace(metadata[1,1], '.+-(.+)', '\\1')))
  base <- paste0(Pasture,"_",Transect)
  sheet_names <- readxl::excel_sheets(path_to_xlsx)[1:3]
  # sheet_names <- readxl::excel_sheets(path_to_xlsx)[2:4]
  # base <- gsub("\\..*","",basename(path_to_xlsx))
  for (j in sheet_names) {
    # j <- sheet_names[2]
    data <- readxl::read_excel(path_to_xlsx, sheet=j)
    if (j == "Metadata") {
      name <- "Meta"
    }else if (j == "Grazed class") {
      name <- "GC"
    }else{
      name <- "SH"
    }
    write.csv(data, file=paste0(finalfolder,"/",base,"_",name,".csv"))
  }
}
```

Make a list of dates for each transect
This will correspond with the day utilization was collected for comparisson to when cattle were moved out of the site. 
```{r, include=FALSE}
(Meta_files <- list.files(path = finalfolder, pattern='Meta',full.names = T))
results <- data.frame(matrix(NA, nrow = length(Meta_files), 
                                 ncol = length(seq(1:2))))
for (i in 1:length(Meta_files)) {
  # i <- "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_10A_Meta.csv"
  metadata <- read.csv(Meta_files[i])
  Pasture <- gsub(" ", "", sub('-.*', '\\1',sub('.*>', '\\1', metadata[1,2])), fixed = TRUE)
  Transect <- gsub(" ", "", gsub(Pasture,"",str_replace(metadata[1,2], '.+-(.+)', '\\1')))
  date <- format(strptime(as.character(gsub(" ","",sub('.*:', '\\1', metadata[2,2]))), "%m/%d/%Y"), "%Y-%m-%d")
  base <- paste0(Pasture,"_",Transect)
  results[i,] <- c(base,date)
  names(results) <- c("Sites", "Date")
  }
write.csv(results, file=paste0(finalfolder,"/",Pasture,"_UseMeasurementdates",".csv"))
```

If there are errors with measruements (measuring in inches instead of cm) you can select the files you need from the "_SH" tagged files and correct the heights in those fodlers only. Then you can rewrite the files afterwrads. 
```{r}
SH<- data.frame(files = list.files(finalfolder, pattern = "SH",full.names = T))
listread <- SH[SH$files %in% c("C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_2_SH.csv",
                   "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_4_SH.csv",
                   "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_7A_SH.csv",
                   "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_13A_SH.csv"), ]
ldf <- lapply(listread, read.csv)

# inch to cm conversion is 2.54
for (i in 1:length(ldf)) {
  ldf[[i]][["Height"]] <- ldf[[i]][["Height"]]*2.54
}
# Names need to be added alphabeltically 
names(ldf) <- c("C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_13A_SH.csv",
                   "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_2_SH.csv",
                   "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_4_SH.csv",
                   "C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/Cleaned_Data/2020/12C/12C_7A_SH.csv") 

for(i in names(ldf)){
  write.csv(ldf[[i]], paste0(i))
}
```

