---
title: "Raw_Export_Combined_Step1"
author: "Brandon Mayer"
date: "4/22/2021"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)  
library(tidyverse)
library(reshape2)

```

Make data frames for each sheet in the excel document

```{r cars}

filename<- ("C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/data/2021/SRER_Use_VGS_Export_RAW_20210816_183142.xlsx")

read_excel_allsheets <- function(filename, tibble = FALSE) {
    # I prefer straight data.frames
    # but if you like tidyverse tibbles (the default with read_excel)
    # then just pass tibble = TRUE
    sheets <- readxl::excel_sheets(filename)
    x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
    if(!tibble) x <- lapply(x, as.data.frame)
    sheets <- gsub(" ","", sheets)
    names(x) <- sheets
    x
}

mysheets <- read_excel_allsheets(filename);rm(read_excel_allsheets)
list2env(mysheets,envir=.GlobalEnv);rm(mysheets);rm(filename)
```
Final product is a rbind df including "GenericValues" and "StubbleHeight" in long form
Need to convert Generic values to long form 

```{r}
CoverValue <- GenericValues[is.na(GenericValues$Value), ]
StubbleValue <- GenericValues[!is.na(GenericValues$Value), ]; rm(GenericValues)

Stubble_gathered <- StubbleValue %>%  rename(V1 = Value, V2 = Value2, V3 = Value3, V4 = CharValue) %>% select(-CharValue2,-CharValue3,-SpeciesQualifier,-FieldQualifier, -SubElement, -Parameter, -Parameter2, -Parameter3,-Element)
Stubble_gathered$V4 <-as.numeric(Stubble_gathered$V4)
Stubble_gathered <-  Stubble_gathered %>%
  pivot_longer(cols = starts_with("V"),
   names_to = "Element", values_to = "Value")
Stubble_gathered$Element[Stubble_gathered$Element == "V1"] <- 000001
Stubble_gathered$Element[Stubble_gathered$Element == "V2"] <- 000002
Stubble_gathered$Element[Stubble_gathered$Element == "V3"] <- 000003
Stubble_gathered$Element[Stubble_gathered$Element == "V4"] <- 000004
Stubble_gathered$Element <-as.numeric(Stubble_gathered$Element)


StubbleHeight<- StubbleHeight %>% select(-SpeciesQualifier,-FieldQualifier, -SubElement, -Parameter, -Parameter2, -Parameter3,-CharValue, -CharValue2, -CharValue3,-Value2, -Value3)
StubbleHeight$Element <-as.numeric(StubbleHeight$Element)
# StubbleHeight$Ancestry <-(StubbleHeight$Ancestry)

head(Stubble_gathered);head(StubbleHeight)
StubbleHeight_comb <- rbind(StubbleHeight, Stubble_gathered);rm(Stubble_gathered, StubbleHeight, StubbleValue)
```
Final product is a rbind df including "Grazed class" and "Item Qualities" in long form

Need to convert "Item Qualities"$values to match grazed classes in "Grazed class" df


```{r pressure, echo=FALSE}
#remove values that dont belong
remove<- setdiff(sort(unique(Grazedclass$SiteID),decreasing = T),sort(unique(CoverValue$SiteID),decreasing = T))

GrazedClass<- Grazedclass %>% filter(!SiteID %in% remove)

Cover_subset <- CoverValue%>% select(Date, SiteID, SampleNumber, "Cover" = CommonName)
# make subsets of cover and same plant 
Sameplant_subset <- Cover_subset %>% filter(Cover %in% "Same Plant") %>% rename("Repeat" = Cover)
Cover_subset <- Cover_subset %>% filter(!Cover %in% "Same Plant")

# combine the subsets and graze class
GrazedClass <- left_join(GrazedClass, Sameplant_subset, by = c("Date","SiteID", "SampleNumber"));rm(Sameplant_subset)
GrazedClass <- left_join(GrazedClass, Cover_subset, by = c("Date","SiteID", "SampleNumber"));rm(Cover_subset)
# replace NA in Repeat with "New plant" 
GrazedClass$Repeat <- GrazedClass$Repeat %>% replace_na("New Plant")

head(GrazedClass);head(CoverValue) # 
```


```{r pressure, echo=FALSE}
#Lets you know if there's any missing graze class values
unique(ItemQualities$Value)
# Identify Graze class values with NAs
ItemQualities[is.na(ItemQualities$Value), ]  
# 
ItemQualities <- ItemQualities[!is.na(ItemQualities$Value), ]  

ItemQualities$Value[ItemQualities$Value == 5] <- 10
ItemQualities$Value[ItemQualities$Value == 20] <- 30
ItemQualities$Value[ItemQualities$Value == 40] <- 50
ItemQualities$Value[ItemQualities$Value == 60] <- 70
ItemQualities$Value[ItemQualities$Value == 80] <- 90

# Use the linking table to correct accessbility standards
linkingtable <- read_excel("data/2021/linkingtable.xlsx")
ItemQualities$Repeat <- linkingtable$SpeciesName[match(ItemQualities$CharValue, linkingtable$PK_Species)]
ItemQualities$Cover <- linkingtable$SpeciesName[match(ItemQualities$CharValue2, linkingtable$PK_Species)]
ItemQualities$CharValue3 <- linkingtable$SpeciesName[match(ItemQualities$CharValue3, linkingtable$PK_Species)]
rm(linkingtable)

# merging Grazeclass and Itemqualities
head(GrazedClass);head(ItemQualities)
merge(x = GrazedClass, y = CoverValue[ , c("Date", "SiteID", "SpeciesName", "CommonName","SampleNumber","Value")], by = c("Date", "SpeciesName", "CommonName","SampleNumber"), all.x=TRUE);rm(CoverValue)


GrazeClass_comb <- rbind(GrazedClass, ItemQualities);rm(GrazedClass, ItemQualities)
GrazeClass_comb <- GrazeClass_comb%>% select(-SpeciesQualifier,-FieldQualifier, -SubElement, -Parameter, -Parameter2, -Parameter3,-CharValue, -CharValue2, -CharValue3,-Value2, -Value3)
GrazeClass_comb$Element <-as.numeric(GrazeClass_comb$Element)

```

Add Pasture column to  
```{r}
StubbleHeight_comb$Pasture<- str_split(StubbleHeight_comb$Ancestry, ' > ', simplify = TRUE)[,4]
GrazeClass_comb$Pasture<- str_split(GrazeClass_comb$Ancestry, ' > ', simplify = TRUE)[,4]
StubbleHeight_comb$Date <- lubridate::mdy(StubbleHeight_comb$Date) 
GrazeClass_comb$Date <- lubridate::mdy(GrazeClass_comb$Date) 
# merge(x = StubbleHeight_comb, y = GrazeClass_comb[ , c("Date", "SpeciesName", "CommonName","SampleNumber","Value")], by = c("Date", "SpeciesName", "CommonName","SampleNumber"), all.x=TRUE)
write.csv(StubbleHeight_comb,"C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/data/2021/stubbleheightSpring.xlsx")
write.csv(GrazeClass_comb,"C:/Users/brand/Dropbox/2.Field_Research/SRER_R_Workshop/data/2021/GrazeClassSpring.xlsx")
```
# Table building
Includes:
1. Average use per transect for joining on ArcGIS
2. Average use for each species per transect
3. Use by pasture
4. Average use for the top three most common species in each pasture

```{r}
path <- paste("C:/Users/brand/Dropbox/Project_Data/SRER/Utilization/Processed_data/", "GY2021/", sep="")
# Percentage Use by site to join with Arcgis site use. 
# colnames(final)[4] <- "Class End-Point"
GrazeClass_comb[is.na(GrazeClass_comb$Cover), ]  
GrazeClass_comb$Cover <- GrazeClass_comb$Cover %>% replace_na("No Cover")
GrazeClass_comb[is.na(GrazeClass_comb$Repeat), ]  
GrazeClass_comb$Repeat <- GrazeClass_comb$Repeat %>% replace_na("New Plant")
final<- GrazeClass_comb

useaverage <- final %>% group_by(Date, Pasture, SiteID) %>%
  summarize(`Percent Ungrazed Utilization` = round(79.9451-(0.8705*sum(Value < 1)), digits = 1),
            `Percent Ungrazed Utilization Repeat` = round(79.9451-(0.8705*sum(Value[Repeat=="New Plant"] < 1)),
                                                           digits = 1),
            `Percent Ungrazed Utilization Accessible` = round(79.9451-(0.8705*sum(Value[Cover=="No Cover"] < 1)),
                                                           digits = 1),
            `Percent Ungrazed Utilization Repeat&Accessible` = round(79.9451-(0.8705*sum(Value[Repeat=="New Plant" & Cover=="No Cover"] < 1)),
                                                           digits = 1),
            `Graze Class Utilization` = round(sum(Value)/sum(Value > 0), digits = 1),
            `Graze Class Utilization Repeat` = round(sum(Value[Repeat=="New Plant"])/sum(Value[Repeat=="New Plant"] > 0), digits = 1),
            `Graze Class Utilization Accessible` = round(sum(Value[Cover=="No Cover"])/sum(Value[Cover=="No Cover"] > 0), digits = 1),
            `Graze Class Utilization Repeat&Accessible` = round(sum(Value[Repeat=="New Plant" & Cover=="No Cover"])/sum(Value[Repeat=="New Plant" & Cover=="No Cover"] > 0), digits = 1)
            )


# write for use in ArcGIS 
write.csv(x = useaverage, file = paste(path, "SiteUse.csv",sep = ""), row.names=FALSE)
```


```{r}
# top species use 
Speciesuseaverage <- final %>% group_by(SiteID, SpeciesName) %>% summarize(count = n(), mean = round(mean(Value)))
Speciesuseaverage <- final %>% group_by(SiteID, SpeciesName) %>% summarize(count = n(), mean = round(mean(Value[Repeat=="New Plant" & Cover=="No Cover"])))

require(data.table)
d <- data.table(Speciesuseaverage, key="SiteID")
d.out <- d[, .SD[mean %in% tail(sort(unique(mean)), 3)], by=SiteID]
d.out <- d.out[order(d.out$SiteID,d.out$mean,decreasing = T)][order(d.out$SiteID,decreasing = F)]
setnames(d.out, old = 'mean', 'Mean Use (%)')
setnames(d.out, old = 'count', 'Percent of Count (%)')
write.csv(x = d.out, file = paste(path, "AllSpeciesUse.csv",sep = ""), row.names=FALSE)
```
```{r}
# Write to append to final document
SiteUseMergeTopSpeciesUse <- left_join(useaverage, d.out, by = c("SiteID"))

UseByTransect <- SiteUseMergeTopSpeciesUse %>% mutate(Transect=mapply(function(x,y)gsub(x,"",y) ,gsub(" ", "|",Pasture),SiteID)) %>%
  select(Pasture, Transect, Date, `Percent Ungrazed Utilization`,"Top Species Used" = SpeciesName, `Percent of Count (%)`, `Mean Use (%)`) 
UseByTransect <- UseByTransect[order(UseByTransect$Pasture,UseByTransect$Transect,decreasing = F),]
write.csv(x = UseByTransect, file = paste(path, "SiteUseByTransect.csv",sep = ""), row.names=FALSE)

```


```{r}
# Percentage of Use by Pasture
pasture_average <- useaverage %>% group_by(Pasture) %>% summarize(year = round(mean(`Percent Ungrazed Utilization`))) 
year <- "GY2021"
data.table::setnames(pasture_average, old = "year", gsub("[/]", "", year))
write.csv(x = pasture_average, file = paste(path, "PastureUse.csv",sep = ""), row.names=FALSE)
```


```{r}
# count, mean - median, and proportion of the use at the site
group_by(Speciesuseaverage,SiteID) %>% group_by(SiteID) %>%
  mutate(pmean=round((count/sum(count))*mean(mean)))

PastureSpeciesUse <- final %>% group_by(Pasture, SpeciesName) %>% summarize(count = n(), mean = round(mean(Value[Repeat=="New Plant" & Cover=="No Cover"]))) %>% mutate(Percent = round(count/sum(count)*100))
P <- data.table(PastureSpeciesUse, key="Pasture")
P.out <- P[, .SD[count %in% tail(sort(unique(count)), 3)], by=Pasture]
P.out <- P.out[order(P.out$Pasture,P.out$count,decreasing = T)][order(P.out$Pasture,decreasing = T)]

setnames(P.out, old = 'mean', 'Mean Use (%)');setnames(P.out, old = 'count', 'Total Count');setnames(P.out, old = 'Percent', 'Percent Count (%)')
P.out
write.csv(x = P.out, file = paste(path, "CommonSpeciesUse.csv",sep = ""), row.names=FALSE)

```




