---
title: "SRER_Utilization"
author: "Brandon Mayer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(readr)
library(googlesheets4)
library(data.table)

```
### Enter the year you're interested in working on
```{r, echo=FALSE}
year <- 'GY2023/' 
```

# Collect original column names to add later 
AND
build tbl to pull meta data from

```{r, include=FALSE}
path <- paste(dirname(path.expand('~')),"/Box/1.Ruyle_lab/1.Project_Data/SRER/Utilization/RAW_Data/", year, sep="")
path2 <- paste(dirname(path.expand('~')),"/Box/1.Ruyle_lab/1.Project_Data/SRER/Utilization/Processed_data/", year, sep="");dir.create(path = paste0(path2))
# files <- list.files(path = path, pattern = "*.csv", full.names = T, all.files = TRUE, recursive = TRUE)
# tbl <- sapply(files, read_csv, simplify=FALSE)
# 
# colnames<- c(colnames(tbl[[1]]))
# rm(files)

```


# Retrieve bulk export files and cleaning
```{r}
sheets<-readxl::excel_sheets(path = list.files(path = path, full.names = T, all.files = TRUE, recursive = TRUE))
filedxlsx <- list.files(path = path, full.names = T, all.files = TRUE, recursive = TRUE) %>%
  lapply(function(x) {readxl::read_excel(x, col_names = TRUE, sheet =sheets[3])})

final <- rbindlist(filedxlsx) %>% select(Date,Ancestry, "Transect"=SiteID,SpeciesSymbol, `Species/Category Name` = SpeciesName, SampleNumber,cValueCommon,cValue2Common,`Class End-Point` = nValue)
final$cValue2Common[is.na(final$cValue2Common)] <- "No Cover"
final$Pasture <- as.character(str_remove_all(str_split_fixed(final$Ancestry, ">",4)[,2],"[ ]"))
final$Date <- lubridate::mdy(final$Date) 
final <- final %>% drop_na(`Class End-Point`) %>% mutate("Month" = lubridate::month(Date),"Year" = lubridate::year(Date))
final <- final %>% filter(Date>=floor_date(max(final$Date), 'year') %m-% months(3)& Date <= max(final$Date))
final <-final %>% mutate(`Class End-Point` = case_when(`Class End-Point`==0~0,
                                               `Class End-Point`==5~10,
                                               `Class End-Point`==20~30,
                                               `Class End-Point`==40~50,
                                               `Class End-Point`==60~70,
                                               `Class End-Point`==80~90))
# filedxlsx <- list.files(path = path, full.names = T, all.files = TRUE, recursive = TRUE) %>%
#   lapply(function(x) {readxl::read_excel(x, col_names = TRUE, sheet =sheets[1])})
# 
# final2 <- rbindlist(filedxlsx) %>% select(Date,Ancestry, "Transect"=SiteID,SpeciesSymbol, `Species/Category Name` = SpeciesName, SampleNumber,cValueCommon,cValue2Common,`Class End-Point` = nValue)
# final2$cValue2Common[is.na(final2$cValue2Common)] <- "No Cover"
# final2$Pasture <- as.character(str_remove_all(str_split_fixed(final2$Ancestry, ">",4)[,2],"[ ]"))
# final2$Date <- lubridate::mdy(final2$Date)
# final2 <- final2 %>% drop_na(`Class End-Point`) %>% mutate("Month" = lubridate::month(Date),"Year" = lubridate::year(Date))
# final2 <- final2 %>% filter(Date>=floor_date(max(final2$Date), 'year') %m-% months(3)& Date <= max(final2$Date))
# 
# final <- rbind(final,final2);rm(final2)
```

# Table building
Includes:
1. Average use per transect for joining on ArcGIS
2. Average use for each species per transect
3. Use by pasture
4. Average use for the top three most common species in each pasture

```{r}
# ---------------------------------------------Percentage Use by site to join with Arcgis site use.----------------------------------------- 
# colnames(final)[4] <- "Class End-Point"
# final$Transect <- str_replace(final$Transect,"D","")
useaverage <- final %>% group_by(Year,Pasture, Month,Transect=as.character(Transect)) %>%
  summarize(`Percent Grazed Utilization (Grazed/Ungrazed)` = round(79.9451-(0.8705*sum(`Class End-Point` < 1)), digits = 1),
            `Graze Class Utilization (Grazed Class)` = round(sum(`Class End-Point`)/n(), digits = 1)) %>% 
  mutate(`Percent Grazed Utilization (Grazed/Ungrazed)` = case_when(`Percent Grazed Utilization (Grazed/Ungrazed)` < 0 ~ 0,
                                               TRUE ~ as.numeric(as.character(`Percent Grazed Utilization (Grazed/Ungrazed)`))))
#useaverage$Transect <- str_replace(useaverage$Transect,"D","")
# useaverage$PastureSite = paste0(useaverage$Pasture, useaverage$Site)

write.csv(x = useaverage, file = paste(path2, "SiteUse.csv",sep = ""), row.names=FALSE)

base<-read_csv(paste(dirname(path.expand('~')),"/Box/1.Ruyle_lab/1.Project_Data/SRER/Utilization/TransectUTM.csv", sep="")) %>% select(-"Shape_1")
useaverage2 <- useaverage
useaverage2$tran_num<- paste0(useaverage2$Pasture,useaverage2$Transect)
useaverage2$tran_num[str_detect(useaverage2$tran_num, "^6E")]<-paste0(useaverage2$tran_num,"_")
useaverage2$tran_num[str_detect(useaverage2$tran_num, "^12E")]<-paste0(useaverage2$tran_num,"_")
base<- base %>% left_join(useaverage2 %>% ungroup() %>% select(tran_num,"GY2022"= `Percent Grazed Utilization (Grazed/Ungrazed)`), by="tran_num")
write.csv(as.data.frame(base),paste0(path2,"TransectU.csv"),row.names = F)
# -------------------------------------------------Average use for each species per transect ------------------------------------------------
Speciesuseaverage <- final %>% group_by(Pasture,Transect, `Species/Category Name`) %>% summarize(count = n(), mean = round(mean(`Class End-Point`)))

d.out <- Speciesuseaverage %>%                                      # Top N highest values by group
  group_by(Pasture,Transect) %>%
  arrange(desc(mean)) %>% 
  slice(1:3)

setnames(d.out, old = 'mean', 'Mean Use (%)')
setnames(d.out, old = 'count', 'Percent of Count (%)')
write.csv(x = d.out, file = paste(path2, "AllSpeciesUse.csv",sep = ""), row.names=FALSE)

# --------------------------------------------------------Percentage of Use by Pasture ------------------------------------------------------
pasture_average <- useaverage %>% group_by(Year,Pasture) %>% summarize(year = round(mean(`Percent Grazed Utilization (Grazed/Ungrazed)`))) 
data.table::setnames(pasture_average, old = "year", gsub("[/]", "", year))
write.csv(x = pasture_average, file = paste(path2, "PastureUse.csv",sep = ""), row.names=FALSE)

# ------------------------------------ Average use for the top three most common species in each pasture ------------------------------------
group_by(Speciesuseaverage,Transect) %>% 
  mutate(pmean=(count/100)*mean(mean))

PastureSpeciesUse <- final %>% group_by(Pasture, `Species/Category Name`) %>% summarize(count = n(), mean = round(mean(`Class End-Point`)))
P <- data.table(PastureSpeciesUse, key="Pasture")
P.out <- P[, .SD[count %in% tail(sort(unique(count)), 3)], by=Pasture]
setnames(P.out, old = 'mean', 'Mean Use (%)')
setnames(P.out, old = 'count', 'Total Count')
P.out
write.csv(x = P.out, file = paste(path2, "CommonSpeciesUse.csv",sep = ""), row.names=FALSE)

```

```{r}
rmarkdown::render(input = paste(dirname(path.expand('~')),"/Box/1.Ruyle_lab/1.Project_Data/SRER/Report Template.RMD", sep=""),
                  output_format = "pdf_document",
                  output_file = paste("SRER",substr(year,1,nchar(year)-1),"Use Report"),
                  output_dir = path2,
                  params = list(
                    year = year,
                  show_code= FALSE
                  )
                  )
```

```{r}
Use <- read_sheet("https://docs.google.com/spreadsheets/d/13UBBNXL4JfQbFNUmtU0sdcR-s1p4nayElb7o1Emh6oQ/edit?usp=sharing")
Use$Pasture <-as.character(Use$Pasture) 
Use$`Transect Name` <-as.character(Use$`Transect Name`)
colnames(Use)[2] <- "Transect"

# Split the df by year, correct naming for months and use, and change negative use values to "0"
uselist <- useaverage %>% split(useaverage$Year)
Y1 <- unique(uselist[[1]]$Year)
Y2 <- unique(uselist[[2]]$Year)
UseEarlySeason <- uselist[[1]] %>% pivot_wider(names_from = Year, values_from = c(Month,`Percent Ungrazed Utilization`)) %>% ungroup() %>% select(-c(`Graze Class Utilization`))
names(UseEarlySeason)[names(UseEarlySeason)%like% "Month"] <- paste0("Month of Measure ",Y1)
names(UseEarlySeason)[names(UseEarlySeason)%like% "Percent Ungrazed"] <- paste0("% Use ",Y1)
# UseEarlySeason[names(UseEarlySeason)%like% paste0("% Use ",Y1)] <- 0

UseLateSeason <- uselist[[2]] %>% pivot_wider(names_from = Year, values_from = c(Month,`Percent Ungrazed Utilization`)) %>% ungroup() %>% select(-c(`Graze Class Utilization`))
names(UseLateSeason)[names(UseLateSeason)%like% "Month"] <- paste0("Month of Measure ",Y2)
names(UseLateSeason)[names(UseLateSeason)%like% "Percent Ungrazed"] <- paste0("% Use ",Y2)
# UseLateSeason[names(UseLateSeason)%like% paste0("% Use ",Y2)] <- 0
```


```{r}
# Add missing columns if necessary
if(!paste0("Month of Measure ",Y1) %in% colnames(Use)){
  Use <- Use %>% add_column(x = c(NA))
  names(Use)[names(Use)%like% "x"] <- paste0("Month of Measure ",Y1)
}
if(!paste0("% Use ",Y1) %in% colnames(Use)){
  Use <- Use %>% add_column(x = c(NA))
  names(Use)[names(Use)%like% "x"] <- paste0("% Use ",Y1)
}
if(!paste0("Month of Measure ",Y2) %in% colnames(Use)){
  Use <- Use %>% add_column(x = c(NA))
  names(Use)[names(Use)%like% "x"] <- paste0("Month of Measure ",Y2)
}
if(!paste0("% Use ",Y2) %in% colnames(Use)){
  Use <- Use %>% add_column(x = c(NA))
  names(Use)[names(Use)%like% "x"] <- paste0("% Use ",Y2)
}

# Add missing values to each column 
for(id in 1:nrow(UseEarlySeason[match(paste(Use$Pasture,Use$Transect),paste(UseEarlySeason$Pasture,UseEarlySeason$Transect)),3])){
  if (is.na(Use[names(Use)%like% paste0("Month of Measure ",Y1)][id,][[1]])) {
    Use[names(Use)%like% paste0("Month of Measure ",Y1)][id,] <- UseEarlySeason[match(paste(Use$Pasture,Use$Transect),paste(UseEarlySeason$Pasture,UseEarlySeason$Transect)),3][id,]
  }
}
for(id in 1:nrow(UseEarlySeason[match(paste(Use$Pasture,Use$Transect),paste(UseEarlySeason$Pasture,UseEarlySeason$Transect)),4])){
  if (is.na(Use[names(Use)%like% paste0("% Use ",Y1)][id,][[1]])) {
    Use[names(Use)%like% paste0("% Use ",Y1)][id,] <- UseEarlySeason[match(paste(Use$Pasture,Use$Transect),paste(UseEarlySeason$Pasture,UseEarlySeason$Transect)),4][id,]
  }
}
for(id in 1:nrow(UseLateSeason[match(paste(Use$Pasture,Use$Transect),paste(UseLateSeason$Pasture,UseLateSeason$Transect)),3])){
  if (is.na(Use[names(Use)%like% paste0("Month of Measure ",Y2)][id,][[1]])) {
    Use[names(Use)%like% paste0("Month of Measure ",Y2)][id,] <- UseLateSeason[match(paste(Use$Pasture,Use$Transect),paste(UseLateSeason$Pasture,UseLateSeason$Transect)),3][id,]
  }
}
for(id in 1:nrow(UseLateSeason[match(paste(Use$Pasture,Use$Transect),paste(UseLateSeason$Pasture,UseLateSeason$Transect)),4])){
  if (is.na(Use[names(Use)%like% paste0("% Use ",Y2)][id,][[1]])) {
    Use[names(Use)%like% paste0("% Use ",Y2)][id,] <- UseLateSeason[match(paste(Use$Pasture,Use$Transect),paste(UseLateSeason$Pasture,UseLateSeason$Transect)),4][id,]
  }
}

# Write the updated df to online google sheet
sheet_write(Use, ss = "https://docs.google.com/spreadsheets/d/13UBBNXL4JfQbFNUmtU0sdcR-s1p4nayElb7o1Emh6oQ/edit?usp=sharing", sheet = "Utilization")

Use <- as.data.frame(Use) 
Use[is.na(Use)] <- ""

xlsx::write.xlsx(Use, paste0(dirname(path.expand('~')),"/Box/1.Ruyle_lab/Santa Rita Experimental Range/Archives/Yearly Utilization.xlsx"), sheetName="Utilization",row.names = F,  append=F)

```

--------------------------------------------------------------Update Info Document----------------------------------------------------

```{r}
#-------------------------------------------------------------Date Updated------------------------------------------------------------

data <- read_sheet("https://docs.google.com/spreadsheets/d/1aJejfLTnhjisjzFHnxK94vw92s276IBTEonQqG6Kvs8/edit?usp=sharing",sheet = "Info")
data[2,1] <- paste0(Sys.Date())
sheet_write(data, ss = "https://docs.google.com/spreadsheets/d/1aJejfLTnhjisjzFHnxK94vw92s276IBTEonQqG6Kvs8/edit?usp=sharing", sheet = "Info")

xlsx::write.xlsx(Use, paste0(dirname(path.expand('~')),"/Box/1.Ruyle_lab/Santa Rita Experimental Range/Archives/percentgrassutilizationstartingin2010.xlsx"), sheetName="Info",row.names = F,  append=F)

#-------------------------------------------------------------Meta data---------------------------------------------------------------
data <- read_sheet("https://docs.google.com/spreadsheets/d/1aJejfLTnhjisjzFHnxK94vw92s276IBTEonQqG6Kvs8/edit?usp=sharing",sheet = "Layout")
y <- str_replace(str_replace(year,"/",""),"GY","")
if (!any(grep(y, data$Item.Name)>1)) {
  
  dataappend <- data.frame("Col." = c(max(data$Col.)+1,max(data$Col.)+2),
    "Item.Name" = c(paste0("Month of Measure ",y),paste0("% Use ",y)),
    "Item.Width" = c(10,10),
    "Data.Type" = c("number","number"),
    "Num.Dec" = c(0,2),
    "Item.Description" = c(paste0("Month of utilization measurement in ",y),paste0("Percent utilization per transect in ",y))
    )
  data <- data %>% rbind(dataappend)
}

sheet_write(data, ss = "https://docs.google.com/spreadsheets/d/1aJejfLTnhjisjzFHnxK94vw92s276IBTEonQqG6Kvs8/edit?usp=sharing", sheet = "Layout")

xlsx::write.xlsx(Use, paste0(dirname(path.expand('~')),"/Box/1.Ruyle_lab/Santa Rita Experimental Range/Archives/percentgrassutilizationstartingin2010.xlsx"), sheetName="Layout",row.names = F,  append=F)
```


